Fast approximate Bayesian inference for small-area estimation of HIV indicators using the Naomi model

### 2023-02-22 src/docs_paper/paper.Rmd
Small paper edits

---
title: Fast approximate Bayesian inference for small-area estimation of HIV using the Naomi model
author:
  - name: Adam 
    surname: Howes
    email: ath19@ic.ac.uk
    label: e1
    addressLabel: A,D
    sepNext: ","
  - name: Alex
    surname: Stringer
    email: alex.stringer@uwaterloo.ca
    label: e2
    addressLabel: B
  - name: Seth R.
    surname: Flaxman
    email: seth.flaxman@cs.ox.ac.uk
    label: e3
    addressLabel: C
    sepNext: ","
  - name: Jeffrey W.
    surname: Eaton
    email: jeffrey.eaton@imperial.ac.uk
    label: e4
    addressLabel: D
affiliation:
  - label: A
    name: Department of Mathematics, Imperial College London
    authorsLabels: e1
  - label: B
    name: Department of Statistics and Actuarial Science, University of Waterloo
    authorsLabels: e2
  - label: C
    name: Department of Computer Science, University of Oxford
    authorsLabels: e3
  - label: D
    name: MRC Centre for Global Infectious Disease Analysis, School of Public Health, Imperial College London
    authorsLabels: e4
abstract: |
  | Naomi is a spatial evidence synthesis model used to produce district-level HIV epidemic indicators in sub-Saharan Africa. Multiple outcomes of policy interest, including HIV prevalence, HIV incidence and antiretroviral therapy treatment coverage, are jointly modelled using both household survey data and routinely reported health system data. Inference for Naomi is currently conducted using an empirical Bayes Gaussian approximation via the \texttt{TMB} \textsc{R} package. We propose a new inference method which combines the simplified integrated nested Laplace approximation approach of @wood2020simplified with adaptive Gauss-Hermite quadrature to enable fast and accurate inference for Naomi and other extended latent Gaussian models. Using data from Malawi, our method provides more accurate inferences than \texttt{TMB}, and is comparable to Hamiltonian Monte Carlo with the No-U-Turn sampler, but faster to run. By extending the \texttt{aghq} \textsc{R} package we facilitate easy, flexible use of our method when provided a \texttt{TMB} \textsf{C++} template for the model's log-posterior. In doing so, we enable inference via integrated nested Laplace approximations for a larger class of models than was previously possible.
keyword-subclass: | 
 \begin{keyword}[class=MSC2020] % It must be define for aap, aop, aos journals. For aoas, sts is not used
 \kwd[Primary ]{00X00}
 \kwd{00X00}
 \kwd[; secondary ]{00X00}
 \end{keyword}
keywords: 
  - spatial statistics
  - small-area estimation
  - INLA
  - AGHQ
  - HIV epidemiology

@@

Mounting an effective public health response to the HIV epidemic requires accurate, timely HIV indicator estimates at the level of geographic resolution at which heath systems are planned and delivered.
Producing these estimates is challenging because all available data sources have shortcomings which must be overcome.
Nationally-representative household surveys provide the most statistically reliable data, but due to their high cost, in most countries they occur only every five years or so, with limited sample size at the district level.
Other data sources, such as routine health surveillance of antenatal care (ANC) clinics, are available in more real-time but based on limited or non-representative samples of the population.
To meet these challenges, the Naomi small-area estimation model [@eaton2021naomi] synthesises data from multiple sources to estimate HIV indicators at a district-level.
Modelling multiple data sources jointly mitigates the limitations of any single source, increases statistical power, and prompts investigation into possibly conflicting information between sources.
Software (\url{https://naomi.unaids.org}) has been developed for Naomi, allowing countries to input their data and interactively generate estimates in a yearly process supported by UNAIDS.
The creation of estimates by country teams, rather than external agencies, is a noteworthy feature of the HIV response.
Drawing on expertise closest to the data being modelled improves the accuracy of the process, as well as strengthening trust and ownership of the estimates.

The practical requirements for the model, combined with its relative complexity, present a difficult Bayesian inference problem.
Any inferential strategy must be fast enough for interactive review and iteration of modelling results, as well as easy to run in production by country teams, ruling out prohibitively slow Markov chain Monte Carlo (MCMC) approaches.
Inference is currently conducted using an empirical Bayes (EB) approach, with a Gaussian approximation to the latent field, via the Template Model Builder (`TMB`) \textsf{R} package [@kristensen2016tmb].
Owing to its speed and flexibility, `TMB` has recently been gaining popularity more broadly in spatial statistics [@osgood2022statistical].
Inference in `TMB` is based on optimisation of a \textsf{C++} template function, with the option available to use a Laplace approximation to integrate out any subset of the function arguments, which for the Naomi model, we use for the high-dimensional latent field parameters.
Taking inspiration from the AD Model Builder (ADMB) package [@fournier2012ad], `TMB` uses automatic differentiation [@baydin2017automatic] to calculate the derivatives required for numerical optimisation routines and the Laplace approximation.
Although this approach has favourable computational properties, it would be preferable to account more fully for hyperparameter uncertainty than is possible in an empirical Bayes framework.
This has motivated us to look for an approach closer to full Bayesian inference, which is flexible enough to be compatible with the model, and fast enough to be run in production by country teams.

To obtain fast, accurate Bayesian inferences for the Naomi model we develop an inference methodology which combines the simplified integrated nested Laplace approximation (INLA) approach of @wood2020simplified with adaptive Gauss-Hermite quadrature (AGHQ).
INLA is an approximate inference approach based on nested Laplace approximations and numerical quadrature.
The key innovation of @rue2009approximate is an approximation which enables accurate latent field posterior marginals without explicitly computing the full Laplace approximation for each element.
Simplified INLA [@wood2020simplified] extends INLA by relaxing the sparsity assumptions on the latent field required for this approximation to be accurate.
This extension facilitates inference for models like Naomi, which are not latent Gaussian models (LGMs) and so were not previously amenable to inference with INLA.
Instead, due to dependence of observations on multiple structured additive predictors, Naomi is what has been termed by @stringer2022fast an extended latent Gaussian model (ELGM).
We combine simplified INLA with AGHQ, a quadrature rule based on the theory of polynomial interpolation which adapts to the integrand based on the Hessian at the mode.
Though no theory yet exists for the nested case, the first stochastic convergence results for adaptive quadrature rules were recently obtained by @bilodeau2021stochastic using AGHQ.
We implement our method as an extension of the `aghq` \textsf{R} package [@stringer2021implementing].
Since `aghq` is designed to naturally interface with `TMB`, use of our method is simple when provided a \textsf{C++} user template for the log-posterior.

@@

@eaton2021naomi specify a joint model linking three small-area estimation models.
The model is defined over three time points: $T_1$ the time of the most recent household survey with HIV testing; $T_2$, the current time period; and $T_3$, a short term projection period.
We consider a simplified version defined only at $T_1$ omitting nowcasting of $T_2$ and temporal projection to $T_3$.
Below we provide an overview of the simplified model, highlighting the aspects which make it a challenge for existing inferential approaches.
A more complete mathematical description of the simplified model, as well as a \textsf{C++} template for the log-posterior, are provided in the appendix.

@@

Let $N_i \in \mathbb{N}$ be the known, fixed population size.
We infer the following unknown HIV indicators: HIV prevalence $\rho_i \in [0, 1]$, the proporiton of indviduals who are HIV positive; antiretroviral therapy (ART) coverage $\alpha_i \in [0, 1]$, the proportion of people living with HIV who receive ART treatment; and annual HIV incidence rate $\lambda_i > 0$, the yearly rate of new HIV infections occurring.
Independent logistic regression models for HIV prevalence and ART coverage in the general population are specified on such that
\begin{align*}
\text{logit}(\rho_i) &= \eta^\rho_i, \\
\text{logit}(\alpha_i) &= \eta^\alpha_i,
\end{align*}
for certain choice of linear predictors $\eta^\rho_i$ and $\eta^\alpha_i$.
For the HIV incidence rate we model on the log scale $\log(\lambda_i) = \eta^\lambda_i(\{\rho_i, \alpha_i\}_{i \in \mathcal{I}})$, where the linear predictor depends on $\{\rho_i, \alpha_i\}_{i \in \mathcal{I}}$ for some $\mathcal{I}$.
Let $\kappa_i$ be the proportion recently infected among HIV positive persons, which we link to HIV incidence via
$$
\kappa_i = 1- \exp \left( - \lambda_i \cdot \frac{1 - \rho_i}{\rho_i} \cdot (\Omega_T - \beta_T) - \beta_T \right),
$$
where the mean duration of recent infection $\Omega_T$ and false recent ratio $\beta_T$ are strongly informed by priors for the particular survey.

For $\theta \in \{\rho, \alpha, \kappa\}$ we calculate the weighted observations
$$
\hat \theta_\mathcal{I} = \frac{\sum_j w_j \cdot\theta_j}{\sum_j w_j},
$$
where $j$ indexes individuals across all strata $i \in \mathcal{I}$.
The design weights are
$$
w_j = \frac{1}{\pi_j} \times \frac{1}{\omega_j},
$$
where $\pi_j$ is the probability of inclusion and $\omega_j$ is a non-response factor for the particular survey.
We calculate the observed number of indicator cases as $y^{\hat \theta}_{\mathcal{I}} = m^{\hat \theta}_{\mathcal{I}} \cdot \hat \theta_{\mathcal{I}}$ where
$$
m^{\hat \theta}_\mathcal{I} = \frac{\left(\sum_j w_j\right)^2}{\sum_j w_j^2},
$$
is the Kish effective sample size [@kish1965survey].

For $\theta \in \{\rho, \alpha, \kappa\}$ we model these aggregate observations using a binomial working likelihood $y^{\hat \theta}_{\mathcal{I}} \sim \text{xBin}(m^{\hat \theta}_{\mathcal{I}}, \theta_{\mathcal{I}})$, where $\theta_{\mathcal{I}}$ are the following appropriately weighted aggregates
\begin{align*}
\rho_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i}{\sum_{i \in \mathcal{I}} N_i}, \\
\alpha_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i \alpha_i}{\sum_{i \in \mathcal{I}} N_i \rho_i}, \\
\kappa_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i \kappa_i}{\sum_{i \in \mathcal{I}} N_i \rho_i}.
\end{align*}

@@

We model HIV prevalence $\rho^\text{ANC}_i$ and ART coverage $\alpha^\text{ANC}_i$ among pregnant women as being offset on the logit scale from the general population indicator as follows
\begin{align*}
\text{logit}(\rho^\text{ANC}_i) &= \text{logit}(\rho_i) + \eta^{\rho^\text{ANC}}_i, \\
\text{logit}(\alpha^\text{ANC}_i) &= \text{logit}(\alpha_i) + \eta^{\rho^\text{ANC}}_i.
\end{align*}
We inform these processes by specifying likelihoods for the following aggregate ANC data from the year of the most recent survey: the number of ANC clients with ascertained status $x^\text{ANC}_\mathcal{I}$, the number of those with positive status $y^\text{ANC}_\mathcal{I}$, and the number of ANC clients already on ART prior to their first ANC visit $z^\text{ANC}_\mathcal{I}$.
We use the binomial working likelihoods
\begin{align*}
y^\text{ANC}_\mathcal{I} &\sim \text{Bin}(y^\text{ANC}_\mathcal{I}, \alpha^\text{ANC}_{\mathcal{I}}), \\
z^\text{ANC}_\mathcal{I} &\sim \text{Bin}(y^\text{ANC}_\mathcal{I}, \alpha^\text{ANC}_{\mathcal{I}}),
\end{align*}
where, again, we use weighted aggregates
\begin{align*}
\rho^\text{ANC}_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} \Psi_i \rho_i^\text{ANC}}{\sum_{i \in \mathcal{I}} \Psi_i}, \\
\alpha^\text{ANC}_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} \Psi_i \rho^\text{ANC} \alpha^\text{ANC}_i}{\sum_{i \in \mathcal{I}} \Psi_i \rho_i^\text{ANC}},
\end{align*}
with $\Psi_i$ the number of pregnant women.

@@

Using one `TMB` template (available in the appendix), we fit the simplified Naomi model to data from Malawi using four inferential approaches: (1) EB combined with a Gaussian approximation using `TMB`, (2) AGHQ combined with a Gaussian approximation using `aghq`, (3) AGHQ combined with a Laplace approximation by extending `aghq` and (4) the Hamiltonian Monte Carlo (HMC) algorithm No-U-Turn Sampling (NUTS) using `tmbstan`.
Details regarding particular settings used for each inferential method are provided in Table \ref{tab:inference-methods}.

>>>>

---
title: Fast approximate Bayesian inference for small-area estimation of HIV indicators using the Naomi model
author:
  - name: Adam 
    surname: Howes
    email: ath19@ic.ac.uk
    label: e1
    addressLabel: A,D
    sepNext: ","
  - name: Alex
    surname: Stringer
    email: alex.stringer@uwaterloo.ca
    label: e2
    addressLabel: B
  - name: Seth R.
    surname: Flaxman
    email: seth.flaxman@cs.ox.ac.uk
    label: e3
    addressLabel: C
    sepNext: ","
  - name: Jeffrey W.
    surname: Eaton
    email: jeffrey.eaton@imperial.ac.uk
    label: e4
    addressLabel: D
affiliation:
  - label: A
    name: Department of Mathematics, Imperial College London
    authorsLabels: e1
  - label: B
    name: Department of Statistics and Actuarial Science, University of Waterloo
    authorsLabels: e2
  - label: C
    name: Department of Computer Science, University of Oxford
    authorsLabels: e3
  - label: D
    name: MRC Centre for Global Infectious Disease Analysis, School of Public Health, Imperial College London
    authorsLabels: e4
abstract: |
  | Naomi is a spatial evidence synthesis model used to produce district-level HIV epidemic indicators in sub-Saharan Africa. Multiple outcomes of policy interest, including HIV prevalence, HIV incidence and antiretroviral therapy treatment coverage, are jointly modelled using both household survey data and routinely reported health system data. Inference for Naomi is currently conducted using an empirical Bayes Gaussian approximation via the \texttt{TMB} \textsc{R} package. We propose a new inference method inspired by adaptive Gauss-Hermite quadrature together with the simplified integrated nested Laplace approximation approach of @wood2020simplified to enable fast and accurate inference for Naomi and other extended latent Gaussian models. Using data from Malawi, our method provides more accurate inferences than \texttt{TMB}, and is comparable to Hamiltonian Monte Carlo with the No-U-Turn sampler, but faster to run. By extending the \texttt{aghq} \textsc{R} package we facilitate easy, flexible use of our method when provided a \texttt{TMB} \textsf{C++} template for the model's log-posterior. In doing so, we enable inference via integrated nested Laplace approximations for a larger class of models than was previously possible.
keyword-subclass: | 
 \begin{keyword}[class=MSC2020] % It must be define for aap, aop, aos journals. For aoas, sts is not used
 \kwd[Primary ]{00X00}
 \kwd{00X00}
 \kwd[; secondary ]{00X00}
 \end{keyword}
keywords: 
  - spatial statistics
  - small-area estimation
  - INLA
  - AGHQ
  - HIV epidemiology

@@

Mounting an effective public health response to the HIV epidemic requires accurate, timely HIV indicator estimates at the level of geographic resolution at which heath systems are planned and delivered.
Producing these estimates is challenging because all available data sources have shortcomings which must be overcome.
Nationally-representative household surveys provide the most statistically reliable data, but due to their high cost, in most countries they occur only every five years or so, with limited sample size at the district level.
Other data sources, such as routine health surveillance of antenatal care (ANC) clinics, are available in more real-time but based on limited or non-representative samples of the population.
To meet these challenges, the Naomi small-area estimation model [@eaton2021naomi] synthesises data from multiple sources to estimate HIV indicators at a district-level.
Modelling multiple data sources jointly has many benefits, including mitigating the limitations of any single source, increasing statistical power, and prompting investigation into any conflicts of information between sources.
Software (\url{https://naomi.unaids.org}) has been developed for Naomi, allowing countries to input their data and interactively generate estimates in a yearly process supported by UNAIDS.
The creation of estimates by country teams, rather than external agencies, is a noteworthy feature of the HIV response.
Drawing on expertise closest to the data being modelled improves the accuracy of the process, as well as strengthening trust and ownership of the estimates.

The practical requirements for the model, combined with its relative complexity, present a difficult Bayesian inference problem.
Any inferential strategy must be fast enough for interactive review and iteration of modelling results, as well as easy to run in production by country teams, ruling out prohibitively slow Markov chain Monte Carlo (MCMC) approaches.
Inference is currently conducted using an empirical Bayes (EB) approach, with a Gaussian approximation to the latent field, via the Template Model Builder (`TMB`) \textsf{R} package [@kristensen2016tmb].
Owing to its speed and flexibility, `TMB` has recently been gaining popularity more broadly in spatial statistics [@osgood2022statistical].
Inference in `TMB` is based on optimisation of a \textsf{C++} template function, with the option available to use a Laplace approximation to integrate out any subset of the function arguments, which for the Naomi model, we use for the high-dimensional latent field parameters.
Taking inspiration from the AD Model Builder (ADMB) package [@fournier2012ad], `TMB` uses automatic differentiation [@baydin2017automatic] to calculate the derivatives required for numerical optimisation routines and the Laplace approximation.
Although this approach has favourable computational properties, it would be preferable to account more fully for hyperparameter uncertainty than is possible in an empirical Bayes framework.
This has motivated us to look for an approach closer to full Bayesian inference, which is also flexible enough to be compatible with the model, as well as fast enough to be run in production by country teams.

To obtain fast, accurate Bayesian inferences for the Naomi model we develop an inference methodology which combines adaptive Gauss-Hermite quadrature (AGHQ) with the simplified integrated nested Laplace approximation (INLA) approach of @wood2020simplified.
INLA is an approximate inference approach based on nested Laplace approximations and numerical quadrature.
The key innovation of @rue2009approximate is an approximation which enables accurate latent field posterior marginals without explicitly computing the full Laplace approximation for each element.
Simplified INLA [@wood2020simplified] extends INLA by relaxing the sparsity assumptions on the latent field required for this approximation to be accurate.
This extension facilitates inference for models like Naomi, which are not latent Gaussian models (LGMs) and so were not previously amenable to inference with INLA.
Instead, due to dependence of observations on multiple structured additive predictors, Naomi is what has been termed by @stringer2022fast an extended latent Gaussian model (ELGM).
We combine simplified INLA with AGHQ, a quadrature rule based on the theory of polynomial interpolation which adapts to the integrand based on the Hessian at the mode.
Though no theory yet exists for the nested case, the first stochastic convergence results for adaptive quadrature rules were recently obtained by @bilodeau2021stochastic using AGHQ.
We implement our method as an extension of the `aghq` \textsf{R} package [@stringer2021implementing].
Since `aghq` is designed to naturally interface with `TMB`, use of our method is simple when provided a \textsf{C++} user template for the log-posterior.

@@

@eaton2021naomi specify a joint model linking three small-area estimation models.
The model is defined over three time points: $T_1$ the time of the most recent household survey with HIV testing; $T_2$, the current time period; and $T_3$, a short term projection period.
We consider a simplified version defined only at $T_1$ omitting nowcasting of $T_2$ and temporal projection to $T_3$.
We provide an overview of the simplified model below, highlighting the aspects which make it a challenge for existing inferential approaches.
A more complete mathematical description of the simplified model, as well as a \textsf{C++} template for the log-posterior, are provided in the appendix.

@@

Let $N_i \in \mathbb{N}$ be the known, fixed population size.
We infer the following unknown HIV indicators: HIV prevalence $\rho_i \in [0, 1]$, the proporiton of indviduals who are HIV positive; antiretroviral therapy (ART) coverage $\alpha_i \in [0, 1]$, the proportion of people living with HIV who receive ART treatment; and annual HIV incidence rate $\lambda_i > 0$, the yearly rate of new HIV infections occurring.
Independent logistic regression models for HIV prevalence and ART coverage in the general population are specified such that
\begin{align*}
\text{logit}(\rho_i) &= \eta^\rho_i, \\
\text{logit}(\alpha_i) &= \eta^\alpha_i,
\end{align*}
for certain choice of linear predictors $\eta^\rho_i$ and $\eta^\alpha_i$.
We model HIV incidence rate on the log scale as $\log(\lambda_i) = \eta^\lambda_i(\{\rho_i, \alpha_i\}_{i \in \mathcal{I}})$, where the linear predictor depends on $\{\rho_i, \alpha_i\}_{i \in \mathcal{I}}$ for some $\mathcal{I}$.
Finally, let $\kappa_i$ be the proportion recently infected among HIV positive persons, which we link to HIV incidence via
$$
\kappa_i = 1- \exp \left( - \lambda_i \cdot \frac{1 - \rho_i}{\rho_i} \cdot (\Omega_T - \beta_T) - \beta_T \right),
$$
where the mean duration of recent infection $\Omega_T$ and false recent ratio $\beta_T$ are strongly informed by priors for the particular survey.

As data, for $\theta \in \{\rho, \alpha, \kappa\}$ we calculate the weighted, aggregated survey observations
$$
\hat \theta_\mathcal{I} = \frac{\sum_j w_j \cdot\theta_j}{\sum_j w_j},
$$
where $j$ indexes individuals across all strata $i \in \mathcal{I}$.
The design weights are given by
$$
w_j = \frac{1}{\pi_j} \times \frac{1}{\omega_j},
$$
where $\pi_j$ is the probability of inclusion and $\omega_j$ is a non-response factor for the particular survey.
We calculate the observed number of indicator cases as $y^{\hat \theta}_{\mathcal{I}} = m^{\hat \theta}_{\mathcal{I}} \cdot \hat \theta_{\mathcal{I}}$ where
$$
m^{\hat \theta}_\mathcal{I} = \frac{\left(\sum_j w_j\right)^2}{\sum_j w_j^2},
$$
is the Kish effective sample size [@kish1965survey].

For $\theta \in \{\rho, \alpha, \kappa\}$ we model these aggregate observations using a binomial working likelihood $y^{\hat \theta}_{\mathcal{I}} \sim \text{xBin}(m^{\hat \theta}_{\mathcal{I}}, \theta_{\mathcal{I}})$, where $\theta_{\mathcal{I}}$ are the following appropriately weighted aggregates
\begin{equation*}
\rho_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i}{\sum_{i \in \mathcal{I}} N_i}, \quad
\alpha_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i \alpha_i}{\sum_{i \in \mathcal{I}} N_i \rho_i}, \quad
\kappa_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i \kappa_i}{\sum_{i \in \mathcal{I}} N_i \rho_i}.
\end{equation*}

@@

We model HIV prevalence $\rho^\text{ANC}_i$ and ART coverage $\alpha^\text{ANC}_i$ among pregnant women as being offset on the logit scale from the general population indicator as follows
\begin{align*}
\text{logit}(\rho^\text{ANC}_i) &= \text{logit}(\rho_i) + \eta^{\rho^\text{ANC}}_i, \\
\text{logit}(\alpha^\text{ANC}_i) &= \text{logit}(\alpha_i) + \eta^{\rho^\text{ANC}}_i.
\end{align*}
We inform these processes by specifying likelihoods for the following aggregate ANC data from the year of the most recent survey: the number of ANC clients with ascertained status $x^\text{ANC}_\mathcal{I}$, the number of those with positive status $y^\text{ANC}_\mathcal{I}$, and the number of ANC clients already on ART prior to their first ANC visit $z^\text{ANC}_\mathcal{I}$.
We use the binomial working likelihoods
\begin{align*}
y^\text{ANC}_\mathcal{I} &\sim \text{Bin}(y^\text{ANC}_\mathcal{I}, \alpha^\text{ANC}_{\mathcal{I}}), \\
z^\text{ANC}_\mathcal{I} &\sim \text{Bin}(y^\text{ANC}_\mathcal{I}, \alpha^\text{ANC}_{\mathcal{I}}),
\end{align*}
where, again, we use weighted aggregates
\begin{equation*}
\rho^\text{ANC}_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} \Psi_i \rho_i^\text{ANC}}{\sum_{i \in \mathcal{I}} \Psi_i}, \quad
\alpha^\text{ANC}_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} \Psi_i \rho^\text{ANC} \alpha^\text{ANC}_i}{\sum_{i \in \mathcal{I}} \Psi_i \rho_i^\text{ANC}},
\end{equation*}
with $\Psi_i$ the number of pregnant women.

@@

Using one `TMB` C++ user-template (available in the appendix), we fit the simplified Naomi model (Section \ref{sec:naomi}) to data from Malawi using four inferential approaches: (1) EB combined with a Gaussian approximation using `TMB`, (2) AGHQ combined with a Gaussian approximation using `aghq`, (3) AGHQ combined with a Laplace approximation by extending `aghq` and (4) the Hamiltonian Monte Carlo (HMC) algorithm No-U-Turn Sampling (NUTS) using `tmbstan`.
Further details regarding particular settings used for each inferential method are provided in Table \ref{tab:inference-methods}.
